---
title: "Web Scrapping with R"
author: "Ayush Patel"
format: html
---

## What will you learn?

These lectures are designed to work as resources for learning the following elements for web scraping:

  1. Identify a desired element on a webpage.
  2. Write R code that will extract the desired element on the web page.
  3. Use Rselenium to automate browsers.


## Who is this for?

if you want to learn how to get information from web pages and are familiar with data wrangling with R. It is likely that you will be able to go through the resources in here with not a copious amount of pain and frustration.

### A little more specific

Take look at the piece of code presented below.

```{r wrangle_check}
#| eval: false

library(palmerpenguins)
library(dplyr)

penguins %>% 
  filter(species == "Gentoo") %>% 
  mutate(
    is_fluffy = ifelse(body_mass_g >= 5000, "Yes","No")
  ) %>% 
  summaries(
    num_fluffy_gentoos = sum(is_fluffy == "Yes", na.rm = T)
  )

```


If you can confidently guess what this code will do on execution or if you have used the functions mentioned in the code in your work, you will be comfortable in using these resources.


## Necessary Set up

### First you will need R and RStudio installed on your machine

  + R version >= 3.2
  + RStudio version >= 2022.7.1.554 (Spotted Wakerobin), though RStudio is not strictly necessary, using this make life easier.
  
If you do not have R installed on your machine, install appropriate version from [here](https://cran.r-project.org/).

if you do not have RStudio installed on your machine, install from [here](https://www.rstudio.com/products/rstudio/).

### Required Libraries

The libraries [Rvest](https://rvest.tidyverse.org/) and [Rselenium](https://docs.ropensci.org/RSelenium/) are required.

```{r libs_install, eval=FALSE}


install.packages("rvest") ## execute to install rvest

install.packages("RSelenium") ## execute to install rselenium

```

You will also need to download the Selenium binary. You can do this from [here](https://selenium-release.storage.googleapis.com/index.html?path=4.0/). Get the file that says `selenium-server-standalone-4.0.0-alpha-1.jar`. Download and execute this file as you would with a .exe file.

Also check if you have java installed on your machine. One way to do this is to execute `java -version` in your cmd or terminal.

## Lecture Slides

[Click this link to navigate to lecture slides](slides/lecuter_slides.html)